### 2025-01-14 paper/content/alg_pto.tex

\todo{elkan's alg. changed to inlcude Pto LBs. too large to include in the paper}
\begin{algorithm}
	\caption{Elkan's Algorithm extended with Ptolemy's inequality}
	\label{alg:ptolemy}

	\begin{algorithmic}

		\State \textbf{Initialization:} Initialize all cluster centers. For each point $x_i$ and each center $c_j$, set the lower bound $l(x_i,c_j)$ and the upper bound $u(x_i)$. Assign each $x_i$ to the nearest cluster $C_j$ such that $c(x_i) = \arg \min_j d(x_i, c_j)$, utilizing Lemma 1 to minimize distance calculations. Set $r(x_i) = \text{true}$ for all points.

		\State \textbf{Step 0:} Compute one iteration of Elkan's algorithm. Save old upper and lower bounds of the last two iterations.
		\Repeat
		\State \textbf{Step 1:}  Compute distances $d(c_i, c_j)$ between all centers, and calculate $s(c_i) = \frac{1}{2} \min_{c_j \neq c_i} d(c_i, c_j)$ for each center $c_i$.
		\Statex
		\State \textbf{Step 2:}  Retain points $x_i$ in their current clusters if $u(x_i) \leq s(c(x_i))$.
		\Statex
		\State \textbf{Step 3:}  For remaining points, consider $x_i$ for reassignment if:
		\begin{itemize}
			\item $c_j \neq c(x_i)$,
			\item $u(x_i) > l(x_i, c_j)$, and
			\item $u(x_i) > \frac{1}{2}d(c(x_i), c_j)$.
		\end{itemize}
		\Statex
		\State \textbf{Step 3a:}  If $r(x_i)$ is true, compute $d(x_i, c(x_i))$. Set $r(x_i) = \text{false}$. Otherwise, $u(x_i) =d(x_i, c(x_i))$
		\Statex
		\State \textbf{Step 3b:} If $d(x_i, c(x_i)) > l(x_i, c_j)$ or $d(x_i, c(x_i)) > \frac{1}{2}d(c(x_i), c_j)$, compute $d(x_i, c_j)$. Reassign $x_i$ to $C_j$ if $d(x_i, c_j) < d(x_i, c(x_i))$.
		\Statex
		\State \textbf{Step 4:} Compute the cluster centers as the centroids of the corresponding clusters $c'_j$.
		\Statex
		\State \textbf{Step 5:} If distance between two older cluster is not 0. Update lower bounds by using equation \ref{eq4.8}. Otherwise, update like in Elkan. Save the old lower bounds of the two last iterations.
		\Statex
		\State \textbf{Step 6:} If distance between two older cluster is not 0. Update upper bounds by using equation \ref{eq4.7}. Otherwise, update like in Elkan. Reset $r(x_i) = \text{true}$. Save old cluster and upper bounds of the last two iterations.
		\Statex
		\State \textbf{Step 7:} Replace each center $c_j$ with $c'_j$.

		\Until{convergence}

	\end{algorithmic}
\end{algorithm}